### 2024-07-29 paper/Cover_letter_Indecisions.tex
supplementary and cover letter added

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% Simple LaTeX CV Template %%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% NOTE: If you find that it says                                     %%
%%                                                                    %%
%%                           1 of ??                                  %%
%%                                                                    %%
%% at the bottom of your first page, this means that the AUX file     %%
%some% was not available when you ran LaTeX on this source. Simply RERUN  %%
%% LaTeX to get the ``??'' replaced with the number of the last page  %%
%% of the document. The AUX file will be generated on the first run   %%
%% of LaTeX and used on the second run to fill in all of the          %%
%% references.                                                        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%% Document Setup %%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Don't like 10pt? Try 11pt or 12pt
\documentclass[12pt]{article}
%\RequirePackage[T1]{fontenc}
\usepackage{ragged2e}
% LaTeX will typeset using Computer Modern Roman, which a lot of
% non-mathematicians and non-engineers won't like. Also, a few PDF
% viewers may not render CMR very well. Instead, Times New Roman can
% be used. That's what this package does.
%times 
%\usepackage{times}
\usepackage{fontspec}

% The automated optical recognition software used to digitize resume
% information works best with fonts that do not have serifs. This
% command uses a sans serif font throughout. Uncomment both lines (or at
% least the second) to restore a Roman font (i.e., a font with serifs).
% (NOTE: This requires the times package above)
%\renewcommand{\familydefault}{\sfdefault}

% This is a helpful package that puts math inside length specifications
\usepackage{calc}

% This package helps LaTeX auto-hyphenate hyphenated words if you use
% special hyphens. For example, bio\-/mimicry will properly hyphenate
% ``mimicry'' if necessary.
\usepackage[shortcuts]{extdash}
\usepackage{graphicx}
\usepackage{fontawesome}
%\usepackage{hyperref}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{graphicx,wrapfig}
%
%         PAPER SIZE, PAGE NUMBER, AND DOCUMENT LAYOUT NOTES:
%
% The next \usepackage line changes the layout for CV style section
% headings as marginal notes. It also sets up the paper size as either
% letter or A4. By default, letter was used. If A4 paper is desired,
% comment out the letterpaper lines and uncomment the a4paper lines.
%
% As you can see, the margin widths and section title widths can be
% easily adjusted.
%
% ALSO: Notice that the includefoot option can be commented OUT in order
% to put the PAGE NUMBER *IN* the bottom margin. This will make the
% effective text area larger.
%
% IF YOU WISH TO REMOVE THE ``of LASTPAGE'' next to each page number,
% see the note about the +LP and -LP lines below. Comment out the +LP
% and uncomment the -LP.
%
% IF YOU WISH TO REMOVE PAGE NUMBERS, be sure that the includefoot line
% is uncommented and ALSO uncomment the \pagestyle{empty} a few lines
% below.
%

%% Use these lines for letter-sized paper
\usepackage[paper=letterpaper,
            includefoot, % Uncomment to put page number above margin
            marginparwidth=0.0in,     % Length of section titles
            marginparsep=.05in,       % Space between titles and text
            margin=1.0in,               % 1 inch margins
            includemp]{geometry}
	   \pagestyle{empty}

\setmainfont{AvenirLTProBook.otf}[
    Path =fonts/,
    BoldFont = AvenirNextLTProBold.otf,
    %BoldFont = AvenirLTProHeavy.otf, % another option for bold font
    ItalicFont = AvenirLTProBookOblique.otf,
    BoldItalicFont = AvenirLTProHeavyOblique.otf
]

%% Use these lines for A4-sized paper
%\usepackage[paper=a4paper,
%            %includefoot, % Uncomment to put page number above margin
%            marginparwidth=30.5mm,    % Length of section titles
%            marginparsep=1.5mm,       % Space between titles and text
%            margin=25mm,              % 25mm margins
%            includemp]{geometry}

%% More layout: Get rid of indenting throughout entire document
\setlength{\parindent}{0in}

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0.0pt} % sets line size of header / footer
\renewcommand{\footrulewidth}{0.0pt}

\usepackage{graphicx} 
\usepackage{caption}
\usepackage{mwe}

\definecolor{myudblue}{RGB}{0, 83, 159}
\usepackage[hidelinks]{hyperref}

%%%%%%%%%%%%%%%%%%%%%%%% End Document Setup %%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% Begin CV Document %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\fontsize{12pt}{12pt}\selectfont
%\pagenumbering{gobble}
\pagestyle{fancy}
\cfoot{}
\lfoot{}
\rfoot{\textcolor{myudblue}{1}}
\raggedright
\begin{minipage}[ht]{0.6\textwidth}
\hspace{-10mm}\includegraphics[height=0.9in]{udcoe_logo}
\end{minipage}\hfill
\begin{minipage}[ht]{0.275\textwidth}
\begin{tabular}{lr}
\raggedleft
&\fontsize{10}{10}\selectfont \hspace{+20mm}\textcolor{myudblue}{Joshua Cashaback}\\
&\fontsize{10}{10}\selectfont \textcolor{myudblue}{Assistant Professor}\\
 &{\fontsize{10}{10}\selectfont \textcolor{myudblue}{\faPhone \hspace{1mm} +1.302.784.5986}}\\
 &{\fontsize{10}{10}\selectfont \textcolor{myudblue}{\href{mailto:joshcash@udel.edu}{\faEnvelope \hspace{1mm} joshcash@udel.edu}}}\\
%&{\fontsize{10}{10}\selectfont \textcolor{myudblue}{\faHome joshcashaback.weebly.com}}
%&{\fontsize{10}{10}\selectfont \textcolor{myudblue}{\href{mailto:cashabackjga@gmail.com}{cashabackjga@gmail.com}}}\\
\end{tabular}
\end{minipage}
\\
\vspace{6mm}
\today
\\
\vspace{6mm}
Dear Editor,
\justify
\vspace{1.0mm}
We are writing to request that you consider our paper, \textcolor{myudblue}{\textbf{``Indecision under time pressure arises from suboptimal switching behaviour''}}, for publication in \emph{Current Biology}.
\vspace{1.0mm}
\\
Indecisions can be catastrophic, preventing a life-saving response or leading to a loss of large sums of money.  Despite their importance and high prevalence, indecisions have been surprisingly ignored in the literature. The main culprit is that the field of decision-making leans heavily on the classic two-alternative forced choice task---where a human or animal \emph{must} select one of two potential options---which simply does not permit a non-response past some deadline (i.e., indecision). Of the few papers that have considered a deadline, indecisions are usually only mentioned in passing and, more often, simply removed from analyses and discarded from model fitting. \textbf{The field has consequently overlooked a common and hallmark feature of human and animal behaviour---indecisions.}
\vspace{3mm}
\\
Here we designed a novel paradigm to study high pressure scenarios that lead to indecisions. We found a median indecision rate of approximately 20\%, \textbf{with some participants even displaying over 50\% indecisions despite having plenty of time to make a guess.} Our computational modelling suggested that these indecisions arose by not accounting for the time delay and uncertainty associated when switching from reacting to guessing. In a second experiment, we show for the first time there is an inherent and functionally relevant time delay when switching from reacting to guessing, a phenomenon previously unreported in the literature. \textbf{Our paper novelly provides a mechanism to explain how indecisions arise.}
\vspace{3mm}
\\
We fully expect that our work will broadly draw interest from fields as diverse as Decision-Making, Economics, and Sports Science, which are all interested in the role of decision-making on behaviour.
Thank you for your time and consideration.
\vspace{3mm}
\\
Sincerely, 
\vspace{3mm}
\\
\includegraphics[height=0.75in]{signature}
\\
\vspace{-3.5mm}
\\
Joshua G. A. Cashaback, Ph.D.
\thispagestyle{fancy}
\rfoot{\textcolor{myudblue}{\thepage}}
\lfoot{\textcolor{myudblue}{\textbf{\fontsize{10}{10}\selectfont Biomedical Engineering | College of Engineering | University of Delaware} }
\\
\fontsize{10}{10}\selectfont \textcolor{myudblue}{STAR Health Sciences Complex, Rm 201J | 540 South College Avenue | Newark, DE., 19713}}
\newpage
\noindent \textcolor{myudblue}{REVIEWER SUGGESTIONS}
\\
\\
Michael S Landy, New York University, landy@nyu.edu
\\
\\
Han L.J. van der Maas, University of Amsterdam, H.L.J.vanderMaas@uva.nl 
\\
\\
Tony Carlsen, University of Ottawa, University of Ottawa
\\
\\
Martin Wiener, George Mason University, mwiener@gmu.edu
%\\
%\\
%Laurence T. Maloney, New York University, laurence.maloney@nyu.edu
%\\
%\\
%Jan Drugowitsch, Harvard Medical School, jan_drugowitsch@hms.harvard.edu
%\\
%\\
%Taraz G. Lee, University of Michigan, tarazlee@umich.edu
%\\
%\\
%Tony Carlsen, University of Ottawa, University of Ottawa
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%% End CV Document %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------%
% The following is copyright and licensing information for
% redistribution of this LaTeX source code; it also includes a liability
% statement. If this source code is not being redistributed to others,
% it may be omitted. It has no effect on the function of the above code.
%----------------------------------------------------------------------%
% Copyright (c) 2007, 2008, 2009, 2010, 2011 by Theodore P. Pavlic
%
% Unless otherwise expressly stated, this work is licensed under the
% Creative Commons Attribution-Noncommercial 3.0 United States License. To
% view a copy of this license, visit
% http://creativecommons.org/licenses/by-nc/3.0/us/ or send a letter to
% Creative Commons, 171 Second Street, Suite 300, San Francisco,
% California, 94105, USA.
%
% THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS
% OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF
% MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT.
% IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY
% CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT,
% TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION WITH THE
% SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
%----------------------------------------------------------------------%